Verifying a Binary Search Tree with Fine-Grained Locking

### 2021-04-13 concurrency/Verification of Concurrent Binary Search Tree.tex
added more content in writting

\usepackage{uri}
\usepackage{mathtools}
\usepackage{amsthm}
%\usepackage{amssymb}
\usepackage{semantic}
\usepackage{graphicx}
\usepackage{cases}
\usepackage{hyperref}
\usepackage{stmaryrd}
\usepackage{listings}
%\usepackage{lstlangcoq}
\renewcommand{\lstlistingname}{Figure}

@@

\newcommand{\TODO}[1]{\textbf{\textcolor{red}{[ TODO: #1]}}}
%\newcommand{\boxdotright}{\!\mathrel\boxdot\joinrel\rightarrow\!}
\newcommand{\islock}{\boxdotright}
\newcommand{\ghost}[2]{#1 @ #2}
\newcommand{\isaex}{\!\mathrel\odot\joinrel\rightarrow\!}
\newcommand{\xisaex}[1]{\!\mathrel\odot\joinrel\xrightarrow{#1}\!}
\newcommand{\ifthenelse}[3]{\text{if }#1\text{ then }#2\text{ else }#3}
\newcommand{\emp}{\mathsf{emp}}

\newcommand{\us}{$\mu$s}

@@

\usepackage{graphicx} % support the \includegraphics command and options
\usepackage{verbatim} % adds environment for commenting out blocks of text & for better verbatim
%\usepackage{amssymb}
\usepackage{semantic}
\usepackage{url}
\usepackage{listings}

@@

\begin{document}
\maketitle

\catcode`\@=11
\section{Introduction}

\section{Background}
\subsection{VST and Iris}
\subsection{Atomic Specifications}

\section{Safety Proofs}
\begin{figure}[htp]
\begin{subfigure}[t]{\textwidth}
\begin{lstlisting}[language = C]
typedef struct tree {int key; void *value; struct tree_t *left, *right;} tree;
typedef struct tree_t {tree *t; lock_t *lock;} tree_t;
typedef struct tree_t **treebox;

void insert (treebox t, int x, void *value) {
  struct tree_t *tgt = *t;
  struct tree *p;
  void *l = tgt>lock;
  acquire(l);
  for(;;) {
    p = tgt->t;
    if (p==NULL) {
      tree_t *p1 = (struct tree_t *) surely_malloc (sizeof *tgt);
      tree_t *p2 = (struct tree_t *) surely_malloc (sizeof *tgt);
      p1 ->t = NULL;
      p2 ->t = NULL;
      lock_t *l1 = (lock_t *) surely_malloc(sizeof(lock_t));
      makelock(l1);
      p1->lock = l1;
      release2(l1);
      lock_t *l2 = (lock_t *) surely_malloc(sizeof(lock_t));
      makelock(l2);
      p2->lock = l2;
      release2(l2);
      p = (struct tree *) surely_malloc (sizeof *p);
      tgt->t = p;
      p->key=x; p->value=value; p->left=p1; p->right=p2;
      release2(l);
      return;
    } else {
      int y = p->key;
      if (x<y){
      	tgt = p->left;
        void *l_old = l;
        l = tgt->lock;
        acquire(l);
        release2(l_old);
      } else if (y<x){
        tgt = p->right;
        void *l_old = l;
        l = tgt->lock;
        acquire(l);
        release2(l_old);
      }else {
      	p->value=value;
        release2(l);
      	return;
      }
    }
  }
} 
\end{lstlisting} 
\end{subfigure}
\caption{Insert Method}
\label{insert}
\end{figure}     
\begin{figure}[htp]
\begin{subfigure}[t]{\textwidth}
 \begin{lstlisting}
void *lookup (treebox t, int x) {
  struct tree *p; void *v;
  struct tree_t *tgt;
  tgt = *t;
  void *l = tgt->lock;
  acquire(l);
  p = tgt->t;
  while (p!=NULL) {
    int y = p->key;
    if (x<y){
      tgt=p->left;
      void *l_old = l;
      l = tgt->lock;
      acquire(l);
      p=tgt->t;
      release2(l_old);
    }else if (y<x){
      tgt=p->right;
      void *l_old = l;
      l = tgt->lock;
      acquire(l);
      p=tgt->t;
      release2(l_old);
    }else {
      v = p->value;
      release2(l);
      return v;
    }
  }
  release2(l);
  return NULL;
}
\end{lstlisting}
\end{subfigure}
\caption{Lookup Method}
\label{lookup}
\end{figure}
In this section, we will focus on safety proofs of concurrent binary search tree. Here safety means thread-safe. An operation of any concurrent data-structure is thread-safe if it functions correctly during simultaneous execution by multiple threads. In VST, we can use $\mathsf{lock\_inv}$ assertion to assert that there exists a lock in memory with a given invariant: $\mathsf{lock\_inv}\ \mathsf{sh}\ p\ R$ means that the current thread owns share $\mathsf{sh}$ of a lock at location $p$ with invariant $R$. We use the term \emph{lock invariant} for $R$, which is a predicate representing the resources protected by a lock. The $\mathsf{lock\_inv}$ predicate is sufficient to prove the safety properties of the concurrent data-structure. We have taken binary search tree with fine-grained concurrent control implemented using hand-over-hand locking mechanism. In hand-over-hand locking mechanism, we acquire the lock for a successor before releasing the lock for a predecessor. The code for $\texttt{insert}$ and $\texttt{lookup}$ methods of concurrent binary search tree that we are interested to verify is presented in Figure 1.

\subsection{Lock invariants for hand-over-hand locking}
A node of CBST have the pointer to its children, so lock invariant of a node should have the knowledge about its children.   
Whenever we release the lock for the parent node, we have to recover all resources that the parent node accessed while acquiring the lock. By releasing the parent node's lock, we would lose the information about the lock\_inv of the node and we wouldn't be able to release the node's lock later. To solve this problem, we need to use \emph{recursive} lock, one whose invariant includes a share of lock itself. In VST, we can make such invariant using $\mathsf{selflock}$ function along with the lemma $\mathsf{selflock\_eq}$: $\forall Q\ \mathit{sh}\ p, \mathsf{selflock}\ Q\ \mathit{sh}\ p = Q * \triangleright \mathsf{lock\_inv}\ \mathit{sh}\ p\ (\mathsf{selflock}\ Q\ \mathit{sh}\ p)$. We can define \emph{lock invariants} and $\mathsf{lock\_inv}$ predicate for \emph{recursive} lock $\mathit{lock}$ as follows: $\mathsf{lock\_inv}\ \mathit{lsh1}\ p\ (\mathsf{selflock}\ P\ \mathit{lsh2}\ \mathit{lock})$ where P represents the knowledge about the children of current node, $\mathit{lsh1}$ and $\mathit{lsh2}$ are the two halves of the writable share $\mathit{Ews}$.

\subsection{Specification and Verification}
\section{Correctness Proofs}
In the previous section, we showed that how to prove concurrent data-structure is thread safe, but not functionally correct. To prove the correctness, our threads need to be able to record the information about the actions they have performed on the shared state, instead of sealing all knowledge of the shared data structure inside the lock invariant. We can accomplish this with ghost variables, a simple form of auxiliary state. Aside from the \emph{permission} on shared resources held by each thread that we used in the safety proofs, the verification of correctness properties also involves the \emph{ghost state}. In VST, any Coq type can be used as ghost state, as long as we can describe what happens when two elements of that type are joined together. We can use own predicate in VST to represent ghost state assertion as follows: $\mathsf{own}\ g\ a\ \mathit{pp}$, where $g$ is a \emph{ghost name}, $a$ is the value associated with $g$ which can be of any type, and $\mathit{pp}$ is a separation logic predicate.

@@

In an Iris-style separation logic, the lock-free approach means that each atomic operation gets access to the state of the data structure stored in an atomic shift, and must either maintain its previous state or (at the linearization point) move to a new one; in the coarse-grained approach we can store the state of the data structure in a lock invariant, and the function has access to all of it throughout the critical section. With fine-grained locking, neither of these approaches quite applies as-is: there are no atomic operations by which we access the state of the data structure, but neither is there a lock whose invariant controls access to the entire structure. However, we can take a modified form of the lock-free approach, where we treat each critical section as an atomic operation that can access the atomic shift. Typically, we will perform a sequence of non-atomic operations in the body of the critical section without access to the global state of the data structure, and then (before exiting the critical section) access the atomic shift and show that we have either maintained the original state or satisfied the postcondition of the operation. In this section, we describe a general approach to associating fine-grained locks with pieces of the abstract state of a data structure, allowing them to interact with atomic shifts and be used to prove atomic specifications. (Note that this approach is not novel to us; it is adapted from something Ralf Jung worked on. We need to be more clear about the relationship.)

>>>>

\usepackage{uri}
\usepackage{mathtools}
\usepackage{amsthm}
%\usepackage{amssymb}
\usepackage{semantic}
\usepackage{graphicx}
\usepackage{cases}
\usepackage{hyperref}
\usepackage{stmaryrd}
\usepackage{listings}
\usepackage{iris}
%\usepackage{lstlangcoq}
\renewcommand{\lstlistingname}{Figure}

@@

\newcommand{\TODO}[1]{\textbf{\textcolor{red}{[ TODO: #1]}}}
%\newcommand{\boxdotright}{\!\mathrel\boxdot\joinrel\rightarrow\!}
\newcommand{\islock}{\boxdotright}
\newcommand{\isaex}{\!\mathrel\odot\joinrel\rightarrow\!}
\newcommand{\xisaex}[1]{\!\mathrel\odot\joinrel\xrightarrow{#1}\!}
%% \newcommand{\ifthenelse}[3]{\text{if }#1\text{ then }#2\text{ else }#3}
\newcommand{\emp}{\mathsf{emp}}

\newcommand\dboxed[1]{\dbox{\ensuremath{#1}}}
\newcommand{\master}[2]{\ensuremath{\mathrm{Master}_{#1}(#2)}}
\newcommand{\snap}[1]{\ensuremath{\mathrm{Snapshot}(#1)}}
\newcommand{\ghost}[2]{\ensuremath{\dboxed{#1}^{#2}}}
\newcommand{\ifthenelse}[3]{\text{if }#1\text{ then }#2\text{ else }#3}
\newcommand{\us}{$\mu$s}

@@

\usepackage{graphicx} % support the \includegraphics command and options
\usepackage{verbatim} % adds environment for commenting out blocks of text & for better verbatim
%\usepackage{amssymb}
\usepackage{url}
\usepackage{listings}

@@

\begin{document}
\acmConference{Some conference}
\maketitle

\catcode`\@=11
\section{Introduction}
Modern CPU design trends have shifted towards multicore processing over the last two decades. This is because the computing power of computers has reached its limit in single-threaded execution, while the quantity of transistors per single-core CPU are increasing as described by Moore's law \cite{591665}. CPUs have multiple cores now because that is the only way to get more transistors per chip, and achieve a desirable throughput. However, doubling the quantity of cores on one chip does not immediately imply an increase in the performance by a factor of two. Programs that we aim to run in multiple cores have to be carefully designed to efficiently divide the workload into threads, and facilitate non-conflicting access to the shared data by multiple threads in parallel. These challenges is resolved by designing the concurrent data structures, algorithms that provide thread-safe access to the shared data by storing and organizing data in a standard way. Concurrent data structures are now widely employed in real-world applications like database algorithms, file systems, networking, e-commerce, etc. Designing these data structures has been considered as one of the difficult tasks in the software development fields. Despite the careful design, implementation, and thorough unit testing of these algorithms, it is very likely to expect subtle bugs and unexpected behaviors of the program. This is because threads interact with each other in complex and subtle ways such that it is almost impossible to cover every corner case while testing. Therefore, it is important to use a precise way of proving the correctness of concurrent data structures which provides a guarantee that the given algorithm correctly implements the specified concurrent data structures.

Formal verification is an active field of research that aims to prove or disprove the correctness of algorithms with respect to a certain formal specification, using mathematical techniques. Model checking and theorem proving  (also called deductive reasoning) are the two main approaches in this field. Model checking builds the mathematical model of a system as a collection of states and actions that move from one state to another. The model checker then outputs yes if the given model meets the given specifications and produces a counterexample otherwise, after a systematic and exhaustive exploration of a mathematical model. Model checking approach only works for finite models and certain infinite models that can be easily abstracted into finite sets of states. While it is a simple and automated operation, it does not scale well to large and complex systems where the search spaces are typically too large to be explored in a reasonable amount of time with a reasonable amount of space. Theorem proving, on the other hand, is a more powerful technique than model checking which can deal with an infinite state space. It works by generating a collection of mathematical proof obligations from the system and its specifications and proving these obligations using either proof assistants (such as Coq, Isabelle, PVS, HOL) or using automatic theorem provers (such as SMT solvers). Theorem proving allows us to work on more accurate representations of the system and express any properties, but most proofs have to be done manually which requires time and expertise. Since the binary search tree that we target for verification is a concurrent data structure of unbounded key-value pairs, and processes. The concurrency considerably increases the state space, and we also want to work with a model that's connected to the actual behavior of C programs, while model checking usually works on more abstract models. So, we use the theorem proving approach to formalize the proof of correctness.

In this thesis, we describe the formal verification of the correctness of concurrent binary search tree with fine-grained locking. We implement the BST algorithms in C and use the Verified Software Toolchain(VST) \cite{10.5555/2670099} to verify with respect to given specification. The concurrent separation logic (CSL) is employed to reason about behavior of the program implemented in the C programming language. Separation logic (SL) \cite{1029817} is an extension of Hoare logic \cite{10.1145/363235.363259}, way of reasoning about programs using a set of logical rules. SL helps to locally reason about programs that play with pointer data structures, where specifications and proofs of a program component only reflect the chunk of memory that the component uses, not the entire global state of the system. Concurrent separation logic (CSL) is a variant of the separation logic for reasoning about the concurrent programs. The version of CSL formalized in VST is rooted on the work of Hobor et al. \cite{10.1007/978-3-540-78739-6_27}, but extended with higher-order ghost state in the style of Iris \cite{10.1145/2951913.2951943}.
\section{Background}
\subsection{VST and Iris}
We use the Verified Software Toolchain \cite{10.5555/2670099}  to prove the correctness of the concurrent binary search tree. VST is a Coq-based proof system that helps to prove separation logic properties of C programs with the assurance that the properties will be preserved in compiled code. First, we write programs in C programming and using CompCert’s \emph{clightgen}, the ASTs for C programs can be generated in Coq. We then state and prove specifications for them using Verifiable C. Verifiable C is a higher-order separation logic for reasoning about the functional correctness of C programs and carries a soundness proof that links high-level specifications to assembly code generated by the CompCert Verified compiler. Verifiable C supports reasoning about two separate kinds of concurrency:  Pthreads-style concurrency with locks (used by implementing blocking data structures that can be fine-grained or coarse-grained ), and concurrency using C11 atomic operations (i.e. lock-free data structures).

\subsection{Ghost State and Global Invariants}
We can prove memory safety and race freedom properties of shared-memory concurrent programs, by tracking the transfer of ownership of shared locations between threads that would happen if we ran the program. But, this idea is insufficient to prove that concurrently running threads are successful to accomplish some task, which is also called functional correctness. For instance, we can prove the memory safety properties of the increment example in  \ref{figure1} through the transfer of ownership of shared variable $x$ between threads. But, we can not guarantee that the value of $x$ will be $2$ after all threads complete their execution, which is demonstrated in \ref{figure1} using separation logic assertion.   
\begin{figure}[htb]
\centering
$\texttt{x = 0;}$\\
$\mathit{I_{\pi}(x \mapsto v \land v \geq 0)}\hspace{30px}\pi = \pi_1\ .\ \pi_2$\\
$\begin{array}{l || l}
\mathit{I_{\pi_1}(x \mapsto v \land v \geq 0)} & \mathit{I_{\pi_2}(x \mapsto v \land v \geq 0)}\\
\texttt{acquire(l);} & \texttt{acquire(l);}\\
\mathit{x \mapsto v \land I_{\pi_1}(x \mapsto v \land v \geq 0)} & \mathit{x \mapsto v \land I_{\pi_2}(x \mapsto v \land v \geq 0)}\\
\texttt{x++;} & \texttt{x++;}\\
\mathit{x \mapsto v+1 \land I_{\pi_1}(x \mapsto v \land v \geq 0)} & \mathit{x \mapsto v+1 \land I_{\pi_2}(x \mapsto v \land v \geq 0)}\\
\texttt{release(l);} & \texttt{release(l);}\\
\mathit{I_{\pi_1}(x \mapsto v \land v \geq 0)} & \mathit{I_{\pi_2}(x \mapsto v \land v \geq 0)}\\
\end{array}$\\
$\mathit{I_{\pi}(x \mapsto v \land v \geq 0)}$
\caption{The increment example annotated with separation logic assertion}
\label{figure1}
\end{figure}
We must preserve the connection between the value of shared resources and the work performed by each thread. A common approach to achieve this is to use \emph{ghost variables}, an auxiliary state introduced later in the proof to track the local information about each thread. They do not appear in the original program. Program in \ref{figure1} can be verified by creating ghost state, which tracks the latest action performed by a thread, for each thread with initial value of $0$, and update with $1$ after each thread increment the value of \texttt{x}. In VST, we can create \emph{ghost state} for any Coq type in the form of an arbitrary \emph{partial commutative monoid} (PCM), a set with a partially defined binary operation that is as associative as it can, commutative, and has a unit, as long as we can describe what happens when two elements of that type are joined together. VST provides $\mathsf {ghost\_var}$ assertion to create a simple ghost state: $\mathsf{ghost\_var}\ \mathit{sh}\ a\ g$ asserts that $g$ is a \emph{ghost name} (\textsf{gname} in Coq) associated with the value $a$, which may be of any type. We will see different kind of ghost states used to verify binary search tree in Chapter \ref{chapter4}.

The main idea behind the logic from Iris \cite{10.1145/3022670.2951943}, a mechnized higher-order concurrent separation logic framework, is the construction of \emph{global invariant} as \emph{ghost state}. \emph{Global Invariant} is an invariant on the global (ghost and physical) state of program. We can open any invariant but must close it again before taking any steps of execution unless those execution are atomic. Thread can use the contents of \emph{global invariant} during atomic operation if it can guarantee that no one will ever see an intermediate state in which invariant does not hold. The rules from Jung et al.\cite{10.1145/3022670.2951943} for creating and opening invariants are:

$$\inference[\textsf{inv\_alloc}]{}{\triangleright P \vdash \pvs[E] \mathsf{EX}\ i : \mathsf{iname}, \knowInv{i}{P}}$$
$$\inference[\textsf{inv\_open}]{i \in E}{\knowInv{i}{P} \vdash \pvs[E][E \setminus i] \triangleright P * (\triangleright P \wand \pvs[E \setminus i][E] \mathsf{emp})}$$

where invariants are provided in the form of assertion  $\knowInv{i}{P}$, and states that $P$ is maintained as an invariant on the global state with name $i$. The operator $\pvs[][]$ is called ``fancy update" operator,that allows to allocate, open, and close the invariants and the later $\triangleright$ operator is used for impredicativity (i.e. $\knowInv{i}{..\knowInv{i}{P}..}$).
\subsection{Atomic Specifications}
\emph{Linearizability} is one of the most popular traditional correctness condition for concurrent data structures. It is defined in terms of a set of high-level operations, such as insert and remove for a set; if a data structure is linearizable, any operation performed on it tends to happen atomically and in some order that corresponds to the real-time ordering of such operations. For instance, if operation $A$ is completed before operation $B$, then $B$ should logically follow $A$. We can prove the linearizability property of concurrent data structures using \emph{global invariants}. Suppose we have an invariant that said ``there exists a linear history of operations performed on the data structures, stored in some ghost state, and the current  state of the data structure is the one that would result from applying each of those operations in order". Then this directly implies that data structure is linearizable. We can allow a data structure operation to access the content of global invariants, treating that operation as an atomic operation if we are able prove that that operation \emph{appears to execute atomically} at some linearization point. This idea was introduced in the TaDA logic \cite{tada} in the form of \emph{atomic Hoare triples} as shown below, and called as  \emph{Logical atomicity}.    
$$\forall a.\ \langle P_l\ |\ P_p(a)\rangle\ c\ \langle Q_l\ |\ Q_p(a)\rangle$$

where $P_l$ and $Q_l$ are \emph{private} pre- and postconditions similar to ordinary Hoare triple, and $P_p$ and $Q_p$ are \emph{public} pre- and postconditions, parameterized by an abstract value $a$ of shared data structure. $P_l$ and $Q_l$ are \emph{local} to each thread 
and do not interfere with global state of a data structure, while $P_p$ and $Q_p$ are public to all threads and must satisfy following rule: $P_p$ must be true at every point from the beginning of the function until the linearization point, for some abstract value $a$, and $Q_p$ must be true at the lineariztion point, but may not still true by the end of the function. So, at the linearization point the operation atomically transitions from $P_p$ to $Q_p$.

VST has encoded an atomic Hoare triple, and provide the way that matches the notation VST uses for normal specifications. Such specification is called \emph{atomic specification}, and can be formalized in Coq as shown in  \ref{atomic_spec}.  This is how we write pre- and postcondition in VST. 
\begin{figure}[htb]
\centering
\begin{verbatim}
Program Definition insert_spec :=
DECLARE _insert
ATOMIC TYPE W OBJ a INVS Ei Eo
WITH ...
PRE [ ... ]
  PROP (...)
  LOCAL (...)
  SEP (P_l) | (P_p)
POST [ ... ]
  PROP ()
  LOCAL ()
  SEP (Q_l) | (Q_p)
\end{verbatim}
\caption{Atomic Specification in VST}
\label{atomic_spec}
\end{figure}
\texttt{W} is the \textsf{TypeTree} representing the type of arguments passed with \texttt{WITH} clause; \texttt{a} is the abstract state for the triple; \texttt{Ei} and \texttt{Eo} are the sets of invariants names inside and outside the triple. The \texttt{PROP} clause describes things that that are true independent of program state, the \texttt{LOCAL} clause describes the values contained in C local variables, and the \texttt{SEP} clause represents the \emph{separating conjunction} (*) of \emph{spatial predicates}, predicates on some part of the memory. In VST, while proving that a function implements an atomic specification, the precondition will contains an $\mathsf{atomic\_shift}$ assertion with the public pre- and postcondition, and the masks inside it. This atomic shift can be accessed through following two rules:
$$\inference[\textsf{atomic\_commit}]{\forall a, R * P_p\ a \Rrightarrow \mathsf{EX}\ y,\ Q_p\ a\ y * R'\ y}{\textsf{atomic\_shift}(P_p, E_i, E_o, Q_p, Q) * R \Rrightarrow \mathsf{EX}\ y,\ Q\ y * R'\ y}$$
$$\inference[\textsf{atomic\_rollback}]{\forall a, R * P_p\ a \Rrightarrow P_p\ a * R'}{\textsf{atomic\_shift}(P_p, E_i, E_o, Q_p, Q) * R \Rrightarrow \textsf{atomic\_shift}(P_p, E_i, E_o, Q_p, Q) * R'}$$
During commit, we must provide resources $R$ which, in combination with the public precondition $P_p$, allow us to prove the public postcondition $Q_p$. Then, we gain access to an assertion $Q$ required by the postcondition of the function, and leftover resources $R'$. During rollback, we provide resources $R$ which, in combination with the public precondition, allow us to reestablish the public precondition; we then regain the atomic shift back in the proof, as well as leftover resources $R'$. The rollback is specially used to learn some relationship between pieces of information (e.g. ghost state) stored in the public precondition, while the commit is used to prove the public postcondition from the precondition, and obtain an assertion $Q$. To complete the proof of any function's specification, we must always perform commit to obtain $Q$; we can perform any number of roolbacks before that point, but after that point we lose the atomic shift and no access to the public precondition.

\section{Safety Proofs}
\label{safety}
\begin{figure}[htp]
\begin{subfigure}[t]{\textwidth}
\begin{lstlisting}[language = C]
typedef struct tree {int key; void *value; struct tree_t *left, *right;} tree;
typedef struct tree_t {tree *t; lock_t *lock;} tree_t;
typedef struct tree_t **treebox;

void insert (treebox t, int x, void *value) {
  struct tree_t *tgt = *t;
  struct tree *p;
  void *l = tgt>lock;
  acquire(l);
  for(;;) {
    p = tgt->t;
    if (p==NULL) {
      tree_t *p1 = (struct tree_t *) surely_malloc (sizeof *tgt);
      tree_t *p2 = (struct tree_t *) surely_malloc (sizeof *tgt);
      p1 ->t = NULL;
      p2 ->t = NULL;
      lock_t *l1 = (lock_t *) surely_malloc(sizeof(lock_t));
      makelock(l1);
      p1->lock = l1;
      release2(l1);
      lock_t *l2 = (lock_t *) surely_malloc(sizeof(lock_t));
      makelock(l2);
      p2->lock = l2;
      release2(l2);
      p = (struct tree *) surely_malloc (sizeof *p);
      tgt->t = p;
      p->key=x; p->value=value; p->left=p1; p->right=p2;
      release2(l);
      return;
    } else {
      int y = p->key;
      if (x<y){
      	tgt = p->left;
        void *l_old = l;
        l = tgt->lock;
        acquire(l);
        release2(l_old);
      } else if (y<x){
        tgt = p->right;
        void *l_old = l;
        l = tgt->lock;
        acquire(l);
        release2(l_old);
      }else {
      	p->value=value;
        release2(l);
      	return;
      }
    }
  }
} 
\end{lstlisting} 
\end{subfigure}
\caption{Insert Method}
\label{insert}
\end{figure}     
\begin{figure}[htp]
\begin{subfigure}[t]{\textwidth}
 \begin{lstlisting}
void *lookup (treebox t, int x) {
  struct tree *p; void *v;
  struct tree_t *tgt;
  tgt = *t;
  void *l = tgt->lock;
  acquire(l);
  p = tgt->t;
  while (p!=NULL) {
    int y = p->key;
    if (x<y){
      tgt=p->left;
      void *l_old = l;
      l = tgt->lock;
      acquire(l);
      p=tgt->t;
      release2(l_old);
    }else if (y<x){
      tgt=p->right;
      void *l_old = l;
      l = tgt->lock;
      acquire(l);
      p=tgt->t;
      release2(l_old);
    }else {
      v = p->value;
      release2(l);
      return v;
    }
  }
  release2(l);
  return NULL;
}
\end{lstlisting}
\end{subfigure}
\caption{Lookup Method}
\label{lookup}
\end{figure}
In this section, we will focus on safety proofs of concurrent binary search tree. Here safety means thread-safe. An operation of any concurrent data-structure is thread-safe if it functions correctly during simultaneous execution by multiple threads. A thread can update the data structure if it holds a sufficiently large share of the data structure so that no other thread can possibly read it. If we want to have a data structure that can be modified by multiple threads, we must move shares between threads via locks. In VST, we can use $\mathsf{lock\_inv}$ assertion to assert that there exists a lock in memory with a given invariant: $\mathsf{lock\_inv}\ \mathsf{sh}\ p\ R$ means that the current thread owns share $\mathsf{sh}$ of a lock at location $p$ with invariant $R$. We use the term \emph{lock invariant} for $R$, which is a predicate representing the resources protected by a lock. The $\mathsf{lock\_inv}$ predicate is sufficient to prove the safety properties of the concurrent data-structure.

%We have taken binary search tree with fine-grained locking. In hand-over-hand locking mechanism, we acquire the lock for a successor before releasing the lock for a predecessor. The code for $\texttt{insert}$ and $\texttt{lookup}$ methods of concurrent binary search tree that we are interested to verify is presented in Figure 1.

\subsection{Lock invariants for hand-over-hand locking}
A node of CBST have the pointer to its children, so lock invariant of a node should have the knowledge about its children.   
Whenever we release the lock for the parent node, we have to recover all resources that the parent node accessed while acquiring the lock. By releasing the parent node's lock, we would lose the information about the lock\_inv of the node and we would not be able to release the node's lock later. To solve this problem, we use \emph{recursive} lock, one whose invariant includes a share of lock itself so that we can retain lock\_inv assertion for a node even after releasing its parent's lock. In VST, we can make such an invariant using $\mathsf{selflock}$ function along with the lemma $\mathsf{selflock\_eq}$: $\forall Q\ \mathit{sh}\ p, \mathsf{selflock}\ Q\ \mathit{sh}\ p = Q * \mathsf{lock\_inv}\ \mathit{sh}\ p\ (\mathsf{selflock}\ Q\ \mathit{sh}\ p)$. We can define $\mathsf{lock\_inv}$ predicate for \emph{recursive} lock $\mathit{lock}$ as follows: $\mathsf{lock\_inv}\ \mathit{sh1}\ p\ (\mathsf{selflock}\ P\ \mathit{sh2}\ \mathit{lock})$ where $\mathit{P}$ represents the knowledge about the children of current node, $\mathit{sh1}$ and $\mathit{sh2}$ are the two halves of the writable share.

\subsection{Specification and Verification}
We begin the verification by defining a predicate $\mathsf{nodebox\_rep}(p)$ that describes the concrete representation of the tree, where $p$ is the pointer to the root node. The $\texttt{nodebox\_rep}$ predicate ties together $\texttt{lock\_inv}$ assertion for each recursive lock as follows: 
\begin{align*}
 &\mathsf{nodebox\_rep}_{\pi}(p) \triangleq  p\mapsto\ (lock,tp)\ *\ \mathsf{lock\_inv_{\pi}}(lock, \mathsf{selflock}_{0.5}(lock,R))) \\&R  \triangleq\ \exists\ pa\ pb\ \ tp\mapsto\ (pa * pb)\ *\  \mathsf{nodebox\_rep}_{0.5}(pa)\ *\ \mathsf{nodebox\_rep}_{0.5}(pb)  \end{align*}

 The ownership of each node's lock is divided into two halves: one half is owned by \emph{lock\ invariant} and another half is owned by node itself. But, in case of root node, we need to split one half into multiple pieces, and share those pieces to all threads such that all of them initially have the knowledge of root node.

\section{Correctness Proofs}
In the previous section, we showed that how to prove concurrent data-structure is thread safe, but not functionally correct. To prove the correctness, our threads need to be able to record the information about the actions they have performed on the shared state, instead of sealing all knowledge of the shared data structure inside the lock invariant. We can accomplish this with ghost variables, a simple form of auxiliary state. Aside from the \emph{permission} on shared resources held by each thread that we used in the safety proofs, the verification of correctness properties also involves the \emph{ghost state}. In VST, any Coq type can be used as ghost state, as long as we can describe what happens when two elements of that type are joined together. We can use own predicate in VST to represent ghost state assertion as follows: $\mathsf{own}\ g\ a\ \mathit{pp}$, where $g$ is a \emph{ghost name}, $a$ is the value associated with $g$ which can be of any type, and $\mathit{pp}$ is a separation logic predicate.

@@

In an Iris-style separation logic, the lock-free approach means that each atomic operation gets access to the state of the data structure stored in an atomic shift, and must either maintain its previous state or (at the linearization point) move to a new one; in the coarse-grained approach we can store the state of the data structure in a lock invariant, and the function has access to all of it throughout the critical section. With fine-grained locking, neither of these approaches quite applies as-is: there are no atomic operations by which we access the state of the data structure, but neither is there a lock whose invariant controls access to the entire structure. However, we can take a modified form of the lock-free approach, where we treat each critical section as an atomic operation that can access the atomic shift. Typically, we will perform a sequence of non-atomic operations in the body of the critical section without access to the global state of the data structure, and then (before exiting the critical section) access the atomic shift and show that we have either maintained the original state or satisfied the postcondition of the operation. In this section, we describe a general approach to associating fine-grained locks with pieces of the abstract state of a data structure, allowing them to interact with atomic shifts and be used to prove atomic specifications. (Note that this approach is not novel to us; it is adapted from something Ralf Jung worked on. We need to be more clear about the relationship.)
 We use the following rules to associate fine-grained locks with abstract state of a data structure:
\begin{align*}
    \textsf{sync\_inv}(g,sh,R) \triangleq  \exists\ a,\ R\ g\ a\ *\ \textsf{my\_half}(g,sh,a)   \end{align*}
$$\inference[\textsf{sync\_commit}]{\forall a, R * P_p\ a \Rrightarrow \exists x_1.\ \mathsf{public\_half}(g, x_1) * \exists x_0'\ x_1'.\ (x_0, x_1) \leadsto (x_0', x_1') \land \\(\mathsf{my\_half}(g, \mathit{sh}, x_0') * \mathsf{public\_half}(g, x_1') \Rrightarrow \exists y.\ Q_p\ a\ y * R'\ y)}{\textsf{atomic\_shift}(P_p, E_i, E_o, Q_p, Q) *  \textsf{sync\_inv}(g,sh,R) \Rrightarrow \mathsf{EX}\ y,\ Q\ y * R'\ y}$$
$$\inference[\textsf{sync\_rollback}]{\forall a, R * P_p\ a \Rrightarrow \exists x_1.\ \mathsf{public\_half}(g, x_1) * (\mathsf{public\_half}(g, x_1) \Rrightarrow P_p\ a * R')}{\begin{array}{c}\textsf{atomic\_shift}(P_p, E_i, E_o, Q_p, Q) * \textsf{sync\_inv}(g,sh,R)  \Rrightarrow \\\textsf{atomic\_shift}(P_p, E_i, E_o, Q_p, Q) * \textsf{sync\_inv}(g,sh,R')\end{array}}$$

The $\mathsf{sync\_inv}$ predicate associates the contribution part of ghost state $a$ for a node with ghost name $g$, along with the other concrete information about a node parameterized on $a$. In the atomic specification, we protect the $\mathsf{sync\_inv}$ using lock inside lock invariant. So, once we acquire the lock, we get access to $\mathsf{sync\_inv}$. The $\mathsf{sync\_commit}$ and $\mathsf{sync\_rollback}$, specialized with ghost state with \emph{reference pattern} ($\mathsf{my\_half}$ and $\mathsf{public\_half}$), are the versions of $\mathsf{atomic\_commit}$ and $\mathsf{atomic\_rollback}$ described in Chapter \ref{chapter2}. In a commit, we must provide $\mathsf{sync\_inv}$ with resources $R$ that, in combination with the public precondition, allow us to prove the public postcondition; we then gain access to an assertion Q required by the postcondition of the function, and leftover resources $R'$. The state of the node $g$ in public precondition (described by an assertion $\mathsf{public\_half}$) must be joinable with the state described by an assertion $\mathsf{sync\_inv}$. In a rollback, we provide $\mathsf{sync\_inv}$ with resources $R$ that, in combination with the public precondition, allow us to reestablish the public precondition; we then regain the atomic shift back in the proof, as well as $\mathsf{sync\_inv}$ with the leftover resources $R$. We must use these rules before releasing the node's lock (we lose the $\mathsf{sync\_inv}$ after releasing the lock).
